Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres

### 2023-08-09 ms.tex
Rogers edits

@@

We use Monte-Carlo simulations of Earth's impact history to determine the most likely timing for the origin of life that requires early RNA synthesis. We use the results of \citet{Wogan_2023}, who find that significant origin of life precursor molecules, such as HCN, are produced in the Hadean atmosphere after a $> 4 \times 10^{20}$ kg impact in an optimistic case, and after a $> 5 \times 10^{21}$ kg impact in a pessimistic scenario. We consider both possibilities, and assume such impacts can cause life's origin so long as they are not followed by a smaller ocean-vaporizing collision that sterilizes the planet.

@@

If life began after an impact that generated a reducing atmosphere, then this work suggests that origin of life on Earth was fairly likely (at least a 20\% chance) and was indeed strongly favored (92\% chance) under optimistic assumptions. Given that rocky planets form from accretionary impacts, our work supports and optimistic outlook for life on exoplanets orbiting sun-like stars and future searches for biosignatures.

@@

N.F.W. and D.C.C. were supported by the Simon's Collaboration on Origin of Life Grant 511570 (to D.C.C.). Also, N.F.W., D.C.C., and K.J.Z. were supported by NASA Astrobiology Program Grant 80NSSC18K0829 and benefited from participation in the NASA Nexus for Exoplanet Systems Science research coordination network. N.F.W. and D.C.C. also acknowledge support from Sloan Foundation Grant G-2021-14194.

>>>>

Furthermore, Hadean zircons provide some constraint on Earth's impact history which we do not explicitly account for in our calculations. The oldest zircons are from Jack Hills, Australia, with U-Pb dates as old as $\sim 4.38$ Ga \citep{Valley_2014}. \citet{Benner_2020} argued that a Moneta-scale impact ($\sim 2 \times 10^{22}$ kg) could not have occurred after this date, because such a collision would heat the crust enough to potentially reset all U-Pb chronometers. \citet{Benner_2020} makes this claim based on \citet{Abramov_2013}, who used models of impact ejecta coupled to simulations of radiogenic Pb-loss in zircons to investigate chronometers resetting during the hypothesized ``late-heavy bombardment''. However, \citet{Abramov_2013} considers impacts much smaller than Moneta (e.g. $< 10^{21}$ kg), and does not clearly establish a minimum impactor mass that would reset all zircon chronometers globally. Given this ambiguity, our simulations do not use zircon chronometers as a constraint on Earth's impact history.

Even if an impact failed to reset a zircon chronometer, the shock wave produced by an asteroid collision may create micro- to nano-structural features in zircons that could be preserved over billions of years \citep{Reimink_2023}. Therefore, the presence or absence of shocked Hadean zircons in the geologic record may provide some constraint on Earth's impact history. For example, \citet{Reimink_2023} estimated the probability of preserved zircons with shock features assuming Earth experienced a ``late-heavy bombardment'' at 3.9 Ga. In this scenario, they find that shocked zircons were likely preserved yet find none in a collection of 4.02 Ga zircons from the Acasta Gneiss Complex. Overall, the absence of preserved shocked zircons suggests that a ``late-heavy bombardment'' did not occur at 3.9 Ga. Our Monte-Carlo models of Earth's impact history do not use zircon shock features as a constraint. A better model would incorporate this information.

@@

We use Monte-Carlo simulations of Earth's impact history to determine the most likely timing for the origin of life that requires early ribonucleobase synthesis. We use the results of \citet{Wogan_2023}, who find that significant origin of life precursor molecules, such as HCN, are produced in the Hadean atmosphere after a $> 4 \times 10^{20}$ kg impact in an optimistic case, and after a $> 5 \times 10^{21}$ kg impact in a pessimistic scenario. We consider both possibilities, and assume such impacts can cause life's origin so long as they are not followed by a smaller ocean-vaporizing collision that sterilizes the planet.

@@

If life began after an impact that generated a reducing atmosphere, then this work suggests that origin of life on Earth was fairly likely (at least a 20\% chance) and was indeed strongly favored (92\% chance) under optimistic assumptions. Given that rocky planets form from accretionary impacts, our work supports an optimistic outlook for life on exoplanets orbiting sun-like stars and future searches for biosignatures.

@@

We thank Roger Buick for reading an early draft of this article and providing useful feedback. N.F.W. and D.C.C. were supported by the Simon's Collaboration on Origin of Life Grant 511570 (to D.C.C.). Also, N.F.W., D.C.C., and K.J.Z. were supported by NASA Astrobiology Program Grant 80NSSC18K0829 and benefited from participation in the NASA Nexus for Exoplanet Systems Science research coordination network. N.F.W. and D.C.C. also acknowledge support from Sloan Foundation Grant G-2021-14194.
